---
title: "Physionet Data Exploration"
author: "David Dai"
date: "Last updated: `r format(Sys.Date(), '%b %d, %Y')`"
output:
    html_document:
        theme: paper
        toc: true
        toc_float: true
        highlight: kate
---


```{r knitr_init, echo=FALSE, cache=FALSE, include = F}
library(knitr)
library(rmdformats)

## Global options
opts_chunk$set(echo=FALSE,
               cache=F,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               cache.lazy = F)
opts_knit$set(width=150)

```

```{r}
library(tidyverse)
library(CHART)
library(here)
library(conflicted)
library(lubridate)
library(mlr)

# Conflicts
conflict_prefer("here", "here")
conflict_prefer("filter", "dplyr")

```

```{r cache = T}
# Work with 1 split for now

train <- read_csv("Z:/LKS-CHART/Projects/physionet_sepsis_project/data/splits/split_0/train_split_0.csv")

test <- read_csv("Z:/LKS-CHART/Projects/physionet_sepsis_project/data/splits/split_0/test_split_0.csv")

data <- bind_rows(train, test)
```

# Summary of the dataset

There are `r nrow(data)` records in the dataset, with `r n_distinct(data$subject)` unique subjects.

## Raw data summary
```{r}
summarizeColumns(data) %>% 
    kable(digits = 2)
```

From the raw summary, a few things are interesting:

1. Numeric ranges are quite extreme in some cases, eg. temperature ranging from 20.9 to 50 
2. No non-missing values of EtCO2
3. Most values are very sparse, with lots of missing data
4. Extremely skewed distributions for various measurements, eg FiO2, AST
5. Gender, Unit1, and Unit2 are categorical variables. Additionally, Unit1 and Unit2 should be mutually exclusive.
    + One of these variables may be sufficient to capture all the information
6. HospAdmTime looks weird
7. Sepsis label is rare, occurring in 2% of records.
8. FiO2 is a percentage; values should fall within 0 to 1.
9. Three variables have negative values: BaseExcess, FiO2, and HospAdmTime. FiO2 should not have negative values; BaseExcess could have negative values; HospAdmTime is the difference in Hospital Admission Time and ICU Admission Time, so a negative value would mean they were admitted to the hospital prior to the ICU.
10. For patients 90 or older, their age was replaced with 100.
11. Each row of data is at an 1-hourly interval.

## Data Pre-processing

Pre-processing steps:

1. Convert gender to factor
2. Convert Unit1, Unit2 to factor
3. Convert SepsisLabel to factor
4. Remove EtCO2 from dataset

```{r}
clean_data <- data %>% 
    mutate(Gender = as.factor(Gender),
           Unit1 = as.factor(Unit1),
           Unit2 = as.factor(Unit2),
           SepsisLabel = as.factor(SepsisLabel)) %>% 
    select(-EtCO2)

```

## Correlations between numeric variables

The figure below shows the Spearman correlations between all numeric variables, using all available pairwise observations. Note that due to high amount of missing data, these correlations may be biased. A deeper understanding of the missing data mechanism is required.

```{r}
cormat <- round(cor(clean_data %>% select_if(is.numeric), use = "pairwise.complete",
                    method = "spearman"),2) 
cormat[upper.tri(cormat)] <- NA

cormat_melted <- reshape2::melt(cormat, na.rm = T)

ggplot(data = cormat_melted, aes(x=Var1, y=Var2, fill = value)) + 
    geom_tile(color = "white") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                         midpoint = 0, limit = c(-1,1),
                         name = "Spearman Correlation") +
    theme_minimal() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## Distribution of numeric variables

A 1% random sample of the data points are plotted below.

```{r}
data %>% 
    sample_frac(0.01) %>% 
    select_if(is.numeric) %>% 
    mutate_all(.f = function(x) ifelse(x==0, 0.001, x)) %>% 
    gather(key = key, value = value) %>% 
    ggplot(aes(y = value, x = key)) + 
    geom_jitter(size = 0.1, alpha = 0.05) +
    coord_flip() +
    # scale_y_continuous(trans = "log10",
    #                    breaks = scales::trans_breaks("log10", function(x)10^x),
    #                    labels = scales::trans_format("log10", scales::math_format(10^.x))) +
    geom_hline(yintercept = 1, lty = "dashed", col = "red") +
    theme_bw()

```



```{r eval = F}
data <- data %>% 
    map(.f = function(x) x %>% 
            mutate(Gender = as.factor(Gender),
                   Unit1 = as.factor(Unit1),
                   Unit2 = as.factor(Unit2),
                   SepsisLabel = as.factor(SepsisLabel)) %>% 
            mutate_if(is.logical, as.numeric))


class_task <- makeClassifTask(id = "classification",
                              data = data[[1]] %>% select(-subject), 
                              target = "SepsisLabel",
                              positive = "1",
)

logreg_learner <- makeLearner("classif.logreg", 
                              id = "logreg",
                              predict.type = "prob",
                              fix.factors.prediction = T)

fit <- train(logreg_learner, class_task)
```